### 2022-09-30 manuscript/manuscript.tex
more rough notes

\begin{align}
  \label{eq:T_2}
  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(t) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
\end{align}

@@

Next, show how quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
essentially model $b(t)$ with an analogous forward in time process. Their
approach is to consider how the variance in neutral allele frequency change is
impacted by a linked selection process. Under a Wright--Fisher model, the
variance in allele frequency change is also determined by a pairwise
coalescence rate, often expressed as a \emph{variance-effective} population
size,

@@

Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
linked selection has no impact on the direction of change, $\E(\delta p_i) =
0$. We derive an expression for $\V(p_t)$ by induction following the approach
of \textcite{Santiago1995-hx}. In the first generation, the frequency change
can be partitioned into three components. Conditioning on the parent's
heritable fitnesses, we can imagine constructing an infinite number of
evolutionary replicates. Across all replicates, a random neutral allele's is
pushed

>>>>

\begin{align}
  \label{eq:T_2}
  \E(T_2) = \sum_{i=1}^\infty \frac{i}{2 b(i) N_d}\prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right).
\end{align}

@@

Next, show how quantitative genetic models of linked selection
\parencite{Robertson1961-ho,Santiago1995-hx,Santiago1998-bs,Santiago2016-mu}
find expressions for $b(t)$ with by modeling the selective process forward in
time. Their approach is to consider how the variance in neutral allele
frequency change is impacted by a linked selection process.

% First, we will show
% how the variance in allele frequency is related to the k

For comparison, we start with the recursion for the variance in allele
frequency change in a neutral Wright--Fisher model. Under binomial sampling of
alleles, the random frequency $x_t$ is $2N x_t \sim \mathrm{Binom}(x_{t-1},
2N)$. If we look forward in time on generation and treat $x_0$ as fixed,

\begin{align}
  \V(x_1) = \E_1\left[(x_1 - x_0)^2\right]= \frac{x_0(1-x_0)}{2N},
\end{align}

and

\begin{align}
  \V(x_2) &= \E\left[(x_2 - x_0)^2\right] \\
          &= \E_1\left[\E_2\left[((x_2-x_1) + (x_1 - x_0))^2 | x_1\right]\right] \\
          &= \E_1\left[\E_2\left[(x_2-x_1)^2|x_1\right]\right] + \E_1\left[ (x_1 - x_0)^2\right] \\
          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N}
\end{align}

such that in general,

\begin{align}
  \label{eq:var_freq}
  %\V(x_t) &= \frac{x_0(1-x_0)}{2N}\left(\sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i-1} \right)
  \V(x_t) &= \V(x_1) \left(1 + \sum_{i=1}^t \left(1 - \frac{1}{2N}\right)^{i} \right)
\end{align}

This geometric series converges to the familiar expression,

\begin{align}
  \V(x_t) &= x_0(1-x_0)\left(1 - \left(1 - \frac{1}{2N}\right)^t \right).
\end{align}

which in the limit $t \to \infty$, converges to the Bernoulli variance
$x_0(1-x_0)$ as the frequency either goes to 1 with probability $x_0$ or 0 with
probability $1-x_0$.

Now, if the effective population size were to change each generation, equation
\eqref{eq:var_freq} would be,

\begin{align}
  \V(x_t) &= \V(x_0) \left(1 + \sum_{i=1}^t \prod_{j=1}^{i} \left(1 - \frac{1}{2N(j)}\right) \right)
\end{align}

Note that this too converges to $x_0(1-x_0)$ as $t \to \infty$.

Note now that equation 4 is a special case, when the evolutionary process has
independent allele frequency changes through the generations. Selection
processes will often lead to correlations in $\delta_i$ and $\delta_j$ for $i
\ne j$. These correlations are because the neutral allele in generations $i$
and $j$ can share some fraction of their fitness background, and in so far as
the selective pressures are the same in generations $i$ and $j$, selection will
perturb the frequency trajectories in the same direction.

One way we can model these dependencies is to partition $\delta_i$ into drift
and selection components. Over evolutionary replicates, the drift components
are independent and the selection components have some degree of
autocorrelation. Consider $\delta x_i = \delta_S x_i + \delta_D x_i$

\begin{align}
  \V(x_2) &= \V\left(\delta x_2 + \delta x_1\right) \\
          &= \V(\delta_D x_2) + \V(\delta_D x_1) + 2 \cov(\delta_S x_2, \delta_S x_1) \\
          &= \frac{x_0(1-x_0)}{2N}\left(1 - \frac{1}{2N}\right) + \frac{x_0(1-x_0)}{2N} +  2 \cov(\delta_S x_2, \delta_S x_1) \\
\end{align}

since the drift changes are the familiar ones of equation XXX. The covariances
are determined by the selective process dynamics, linkage disequilibrium, and
the recombination rate.

Let us imagine there is some stochastic process $b(t)$ that rescales the coalescent
rate,

\begin{align}
  \V(x_t) &= \frac{x_0(1-x_0)}{2N(0)}\left(\sum_{i=1}^t \prod_{j=1}^{i-1} \left(1 - \frac{1}{2b(j)N}\right) \right)
\end{align}

\begin{align}
  \var(p_3 - p_0) = \E & \left( \bigg( \underbrace{S_1(1 + (1-r) G + (1-r)^2 G^2)}_{\text{creation and persistence of generation 1 associations} \;\; := \; S_1 Q_3} \right. +  D_1 + H_1+ \nonumber \\
                       & \underbrace{S_2(1 + (1-r) G)}_{\text{creation and persistence of generation 2 associations} \;\; := \; S_2 Q_2}+ D_2 + H_2 + \nonumber \\
                       &  \underbrace{S_3}_{\text{creation of generation 3 associations} \;\; := \; S_3 Q_1  }\left. + D_3 + H_3 \bigg)^2 \right). &
    \label{eq:scr-var2}
\end{align}
%

Note the similarity between the equations.

\begin{align}
  V &\propto \frac{1}{2b(0)N_d}\left(\sum_{i=1}^\infty \prod_{t=1}^{i-1} \left(1 - \frac{1}{2b(t)N_d}\right) \right) \\
  W &\propto \sum_{i=1}^\infty \frac{i}{2 b(i) N_d} \prod_{t=1}^{i-1} \left(1-\frac{1}{2 b(t) N_d}\right) \\
\end{align}

Let us define the function $a(t)$ to be the product, which is the cumulative
failure to coalesce up to time $t$.

\begin{align}
  a(t) = \prod_{i=1}^{t-1} \left( 1-\frac{1}{2 b(i) N_d} \right)
\end{align}

Comparing this to equation \eqref{eq:T_2} up to constant factors,

\begin{align}
  V &\propto \sum_{i=1}^\infty a(t) \\
  W &\propto \sum_{i=1}^\infty i \; a(t) \\
\end{align}

end{align}

XXX we swap indices

Now, if we assume that the final coalesce in the waiting time process were to
have a fixed rate of $\nicefrac{1}{2b(0)N_d}$, this

we see that both the waiting time process and the variance process

Intuitively, these define a stochastic selection process in two ways through
the function $b(t)$ that modulates the rate of pairwise coalesce.

@@

Since we track an arbitrary neutral allele's frequency $p_i$, by symmetry,
linked selection has no impact on the direction of change, $\E(\Delta p_i) =
0$. We will work through a short proof of $\V(p_t)$ by induction following the
approach of \textcite{Santiago1995-hx}. Throughout, we will use angular
brackets to indicate an average over all diploid individuals in the population,
of some value with an individual-level index. For example, if we denote the
number of allele an individual carries in generation $t$ as $x_{t,i} \in \{0,
1, 2\}$, the population frequency at time $p_t$ is

\begin{align}
  p_t = \langle x_t \rangle = \frac{1}{N} \sum_{i=1}^N x_{t,i}
\end{align}

\begin{align}
  \Delta_H p &= \langle x (f-1) \rangle \\
  \Delta_N p &= \langle x d \rangle \\
  \Delta_M p &= \langle \delta \rangle \\
\end{align}

Assuming uninherited, random environments, only the $\Delta_{_H} p$ terms
create autocorrelation in $p_t$. The autocorrelation is created because during
the random pairing of neutral alleles to fitness backgrounds, associations are
created between an individual $i$'s allele frequency and their fitness $f_i$.
This is because both fitness and the neutral alleles reside in the same
individual in the present generation.

We can imagine now how these associations change over evolutionary replicates
in the next generation. In a fraction $r$ of the cases, the neutral allele 
recombines off (or segregates away, if they are on different chromosomes) the
fitness background. In a faction of $1-r$ XXX

\begin{align}
  \cov(\Delta_H p_t, \Delta_H p_s) &= \E(\langle x_t (f_t-1) \rangle , \langle x_s (f_s-1) \rangle)
\end{align}

Note that this covariance is only

In the first generation, the frequency change
can be partitioned into three components. Conditioning on the parent's
heritable fitnesses, we can imagine constructing an infinite number of
evolutionary replicates. Across all replicates, a random neutral allele's is
perturbed
